---
title: "bmes547 hw3"
author: "Jeremy Leipzig"
date: "2/8/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Classification of Active Sites using Neural Network
In this assignment, you are going to predict catalytic residues in proteins using sequence and structural information. The dataset (courtesy of Natalia Petrova) is a subset of the data used in "Prediction of catalytic residues using Support Vector Machine with selected protein sequence and structural properties", Natalia Petrova and Cathy Wu, 2006.

http://bmcbioinformatics.biomedcentral.com/articles/10.1186/1471-2105-7-312

```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(nnet)
library(caret)
library(dplyr)
library(stringr)
library(ggplot2)

set.seed(1234)
```

```{r, echo=TRUE, warning=FALSE, message=FALSE, error=FALSE, results='hide'}
## Load the data
# Get the data using the loadcatsite() fnction.
source("loadcatsite.R")
X<-loadcatsite()$X
T<-loadcatsite()$T

#create sets on the targets to insure even distribution of outcomes
flds <- createFolds(T, k = 4, list = TRUE, returnTrain = FALSE)

test.cl <- function(true, pred) {
  #return confusion matrix
  truef <- factor(colnames(true)[max.col(true)],levels=c(-1,1))
  predf <- factor(colnames(pred)[max.col(pred)],levels=c(-1,1))
  ct<-table(truef, predf)
  return(ct)
}

get_accuracy<-function(resulttable){
  #assumes NEG POS columns and rows
  #Classification accuracy
  #(TP + TN) / (TP + TN + FP + FN)
  TN<-resulttable[1,1]
  TP<-resulttable[2,2]
  FN<-resulttable[2,1]
  FP<-resulttable[1,2]
  accuracy <- ((TP + TN) / (TP + TN + FP + FN))
  return(accuracy)
}



max.units<-30
num.folds<-4

training_classification_accuracy<-NULL
test_classification_accuracy<-NULL

training_fold<-list()
test_fold<-list()

for(units in 1:max.units){
  print(paste(units,"units"))
  training_fold[[units]]<-list()
  test_fold[[units]]<-list()
  for(fold in 1:num.folds){
    trainingset <- unlist(flds[-fold])
    testset <- flds[[fold]]
    net<-nnet(X[trainingset,],class.ind(T[trainingset]),size = units, maxit = 500)
    training_fold[[units]][[fold]]<-test.cl(class.ind(T[trainingset]), predict(net, X[trainingset,]))
    test_fold[[units]][[fold]]<-test.cl(class.ind(T[testset]), predict(net, X[testset,]))
  }
  training_classification_accuracy[units]<-get_accuracy(Reduce("+",training_fold[[units]]))
  test_classification_accuracy[units]<-get_accuracy(Reduce("+",test_fold[[units]]))
}
data.frame(units=1:max.units,training_classification_accuracy,test_classification_accuracy) -> results
```

## What is the training and test accuracies?
```{r}
results %>% melt(id.vars="units") -> melt_results

ggplot(melt_results,aes(units,value))+
  geom_smooth(aes(color=variable))+
  geom_point(aes(color=variable))+
  xlab("hidden units")+
  ylab("classification accuracy")

knitr::kable(results %>% select(units,test_classification_accuracy))
```

## What is the best number of hidden units?
`r results$units[which(results$test_classification_accuracy==max(results$test_classification_accuracy))]` units


