---
title: "bmes547 hw1"
author: "Jeremy Leipzig"
date: "1/25/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(ggplot2)
library("MASS")
set.seed(1)
```

## Q1
```{r}
# This function simulates flipping a coin by generating random numbers 
# from a uniform distribution in [0,1) and considering numbers < 0.5 to be 
# tails and numbers >= 0.5 heads.
# Returns the ratio of the heads over "trials" number of coin flips.
hw1_coinflips<-function(trials){
 flip = runif(trials)
 heads = (flip >= 0.5);
 headsratio = sum(heads)/trials;
 return(headsratio)
}

```


### a
```{r, echo=TRUE, message=TRUE, warning=FALSE, results='asis'}
print(hw1_coinflips(100))
hw1_coinflips(100)
```
No, the answer is not the same because of sampling error.

### b
```{r, echo=TRUE, message=TRUE, warning=FALSE, results='asis'}
getsd<-function(trials,experiments){
 headsratios<-sapply(1:experiments,function(x){hw1_coinflips(trials)})
 return(sd(headsratios))
}
trials<-c(10,100,1000,10000,100000)
results<-data.frame(trials=trials,sd_trials=sapply(trials,function(x){getsd(x,10)}))
ggplot(results,aes(trials,log(sd_trials)))+geom_point()
```

The standard deviation of a binomial distribution with N trials and a ${\mu}$ probability of success on each trial is:

$\sqrt {N\mu(1-\mu)}$
for ${\mu}$ of .5 this is
$\sqrt {N*.5(.5)}$


### c
```{r, echo=TRUE, message=TRUE, warning=FALSE, results='asis'}
experiments<-c(10,100,1000,10000,100000)
results<-data.frame(experiments=experiments,sd_exp=sapply(experiments,function(x){getsd(10,x)}))
ggplot(results,aes(experiments,log(sd_exp)))+geom_point()
```

We would expect the standard deviation to increase with the number of trials, since if you flip a coin but the number of trials is constant here. Increasing the number of experiments reduces the sampling error so we are just asymptotically approaching at `r sqrt(10*.25)`, the log of which is `r log(sqrt(10*.25))`

\pagebreak

## Q2
```{r}
hw1_montyhall<-function(strategy,trials){
  sample.vec <- function(x, ...) x[sample(length(x), ...)]
  wins<-0
  for(i in 1:trials){
    #randomly determine which door has the car.
    cardoor <- sample.vec(1:3,1)
    #player picks a door randomly.
    playerdoor <- sample.vec(1:3,1)
    #among the two doors the player did not pick,
    #randomly discard one that doesn't have the car.
    discarddoor<-sample.vec((1:3)[c(-cardoor,-playerdoor)],1)
    if(strategy == 1){
      playerdoor<-(1:3)[c(-discarddoor,-playerdoor)]
    }else{
      if(strategy == 2){
        playerdoor<-sample.vec((1:3)[c(-discarddoor)],1)
      }
    }
    #strategy 0 do nothing
    if(playerdoor==cardoor){wins<-wins+1}
  }
  return(wins/trials)
}
```
never switch door
```{r}
hw1_montyhall(0,1000)
```

always switch door
```{r}
hw1_montyhall(1,1000)
```

switch door at random
```{r}
hw1_montyhall(2,1000)
```
\pagebreak

## Q3
```
P(alarm+) = 0.9
P(alarm-) = 1 - P(alarm+) = 0.1
P(getup|alarm+) = 0.95
P(getup|alarm-) = 0.60
P(getup) = P(getup|alarm+)*P(alarm+) + P(getup|alarm-)*P(alarm-) = .95(0.9) + .60(0.10) = 0.915
```

## Q4
```{r}
generate_uniform<-function(i){
  n<-1000
  return(sum(runif(n))-n*0.5)
}
all_samples<-sapply(1:1000,generate_uniform)

expected_variance<-1/(.7234*sqrt(1000))
expected_sd<-sqrt(expected_variance)

observed_sd<-sd(all_samples)

#we need to scale up the density to plot this on a histogram
ggplot() + aes(all_samples)+geom_histogram()+stat_function( 
  fun = function(x, mean, sd, n, bw){ 
    dnorm(x = x, mean = mean, sd = sd) * n * bw
  }, args = c(mean = 0, sd = observed_sd, n = 2250, bw = 1))
```
\pagebreak

## q5
```{r}
X<-rnorm(1000)

ggplot()+aes(X)+geom_histogram(bins=100)
# a tabulation of the counts in each bin
# for clarity's sake let's call this Y
Y<-table(cut_interval(X,n=100))
Y<-Y/sum(Y)
muY<-mean(Y)
varianceY<-var(Y)
muY
varianceY

sigmaY<-sqrt(varianceY)


ggplot(data.frame(x = c(-0.1, 0.1)), aes(x)) + 
  geom_density() + 
  stat_function(fun = dnorm, args = list(mean = muY, sd = sigmaY))


fit<-fitdistr(Y, "normal")
fit$estimate
```
Yes, the mean and standard deviation look similar to the one from the tabulation
\pagebreak

## q6
```{r}
x_given_w1<-function(x){ dnorm(x, mean = 1, sd = 1) }
x_given_w2<-function(x){ dnorm(x, mean = 2, sd = 4) }

ggplot(data.frame(x = c(-10, 10)), aes(x)) + 
  geom_density() + 
  stat_function(fun = x_given_w1, aes(color = "x|w1")) +
  stat_function(fun = x_given_w2, aes(color = "x|w2")) + 
  ylab("Class densities") +
  scale_colour_manual("Legend", values = c("red","blue"))

P_w1 <- 0.4
P_w2 <- 0.6
P_x <- function(x){x_given_w1(x) + x_given_w2(x)}
w1_given_x<-function(x){(x_given_w1(x)*P_w1)/P_x(x)}
w2_given_x<-function(x){(x_given_w2(x)*P_w2)/P_x(x)}
ratio<-function(x){(x_given_w1(x)*P_w1)/(x_given_w2(x)*P_w2)}
decision_threshold_function <- function(x){abs(ratio(x)-1)}
decision_threshold_one<-optimize(decision_threshold_function,interval = c(-2,0))
decision_threshold_two<-optimize(decision_threshold_function,interval = c(2,5))

ggplot(data.frame(x = c(-10, 10)), aes(x)) +
  stat_function(fun = w1_given_x,geom = "line", aes(color = "w1|x")) +
  stat_function(fun = w2_given_x,geom = "line", aes(color = "w2|x")) +
  stat_function(fun = ratio, geom = "line", aes(color= "ratio")) + ylab("Posterior Probabilities") +
  geom_vline(xintercept = decision_threshold_one$minimum) +
  geom_vline(xintercept = decision_threshold_two$minimum) +
  scale_colour_manual("Legend", values = c("green","red","blue"))

```

w1 is more likely if x is 1